---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

  
### Loading and preprocessing the data
```{r, echo = FALSE}
setwd("C:/users/btaylor/desktop/reproducible_research/")
```
```{r load_data, message = FALSE}
##setwd(...)

library(lubridate)
library(dplyr)
library(ggplot2)

df <- read.csv("activity.csv", stringsAsFactors = FALSE)
df$date <- ymd(df$date)
df$week_period <- factor(
   ifelse(weekdays(df$date) %in% c("Saturday", "Sunday"), "Weekend", "Weekday"))
```
  
### What is mean total number of steps taken per day?
```{r simple_stats}
x <- df %>% 
  group_by(date) %>% 
  summarise(total = sum(steps, na.rm = TRUE))

ggplot(x) + 
  geom_histogram(aes(x = total), fill = "blue", alpha = 0.5, bins = 20) + 
  labs(x = "", y = "Frequency", title = "Total Steps per Day") +
  theme_bw()
```
  
The mean total steps taken per day is `r prettyNum(round(mean(x$total)), big.mark = ",")`.  
The corresponding median is `r prettyNum(round(median(x$total)), big.mark = ",")`.

  
### What is the average daily activity pattern?
```{r avg_daily}
x <- df %>% 
  group_by(interval) %>%
  summarise(avg_steps = mean(steps, na.rm = TRUE))

ggplot(x, aes(x = interval, y = avg_steps)) + 
  geom_line(color = "blue", alpha = 0.5) +
  labs(x = "", y = "", title = "Average Daily Steps by 5-Min. Interval") +
  theme_bw()
```
  
Interval with maximum average value is `r x$interval[which.max(x$avg_steps)]`.
  
### Imputing missing values
```{r n_missing}
## Percent of missing values
n_missing <- prettyNum(sum(is.na(df$steps)), big.mark = ",")
pct_missing <- paste0(round(sum(is.na(df$steps)) / nrow(df) * 100, 2), "%")
```
Number of step observations missing = `r n_missing`.  
Percent of step observations missing = `r pct_missing`.

```{r impute missing}
## Methodolgy:
## 1. Calculate average value by interval
## 2. Replace missing values with average interval value
df_all <- df %>%
  group_by(interval) %>%
  mutate(avg_steps = mean(steps, na.rm = TRUE),
         steps = ifelse(is.na(steps), avg_steps, steps),
         avg_steps = NULL)

## Daily pattern data for plotting
x <- df_all %>%
  group_by(interval) %>%
  summarise(total = sum(steps))

ggplot(x) + 
  geom_histogram(aes(x = total), fill = "blue", alpha = 0.5, bins = 20) + 
  labs(x = "", y = "Frequency", title = "Total Steps per Day\nWith Imputed Missing Values") +
  theme_bw()
```
  
**With imputed missing values**

The mean total steps taken per day is `r prettyNum(round(mean(x$total)), big.mark = ",")`.  
The corresponding median is `r prettyNum(round(median(x$total)), big.mark = ",")`.
  
Each statistics is much lower than those calculated on the dataset without imputed missing values.


### Are there differences in activity patterns between weekdays and weekends?
```{r week_period}
x <- df_all %>%
  group_by(week_period, interval) %>%
  summarise(avg_steps = mean(steps))

ggplot(x) +
  geom_line(aes(x = interval, y = avg_steps), color = "blue", alpha = 0.5) +
  facet_wrap(~week_period, nrow = 2) +
  labs(x = "", y = "", title = "Average Daily Steps by 5-Min Interval\nWith Imputed Missing Data") +
  theme_bw()


```


